---
title: "BCest_simulation"
output: html_document
---


## Simulation Process

We simulate covariates $X_{1}$ independently from normal distribution with known mean and variance to generate the covariate matrix $X$,
$$
X = 
\begin{bmatrix}
    1 & x_{11}\\
    1 & x_{12} \\
    \vdots \\
    1 & x_{1n} 
\end{bmatrix}
$$

To generate the response matrix Y, we use some true underlying model $Y = g(X) + \epsilon$. For example, we set $g(X)$ to be a linear regression model $g(X) = X\beta$, where X is the simulated covariate matrix, $\beta$ is given, and $\epsilon$ is also simulated.

```{r eval=FALSE}
library(dplyr)
library(caret)
library(parallel)
```

We want to split our covariates and responses randomly into training, testing, and validation data sets in proportion 6:2:2, so we first sample data randomly into 10 sub-groups. In the `creat_testIndex(sample_size)` function, it takes the sample size as the parameter, and randomly samples sample size into 10 sub-groups.
```{r eval=FALSE}
create_testIndex <- function(sample_size){
  Index_list <- list()
  group_idx <- sample(cut(seq(1,sample_size),breaks=10,label=FALSE))
  for (i in 1:10){
    Index_list[[i]] <- which(group_idx == i)
  }

  return(Index_list)
}

```


In the `simulation(sample_size, num_of_cov, beta_list, Index_list, use_seed=12345)` function, it takes in 5 parameters: sample size, number of covariates, a list of given betas, an index list created from `creat_testIndex(sample_size)`, and a seed number for simulation purpose.
```{r eval=FALSE}
simulation <- function(sample_size, num_of_cov, beta_list, Index_list, use_seed=12345){

  set.seed(use_seed)
```

We create a list of covariates `covariate_list`, where the first element in the list is a vector of 1s where the length is the same as the sample size. The second element of `covariate_list` is $X_1$, where we simulate from $\mathcal{N}(0,1)$ (assuming we have standardized covariates) and the length of $X_1$ is equal to the sample size.

```{r eval=FALSE}
  covariate_list <- list()
  covariate_list[[1]] <- rep(1, sample_size)
  for (i in c(2:(num_of_cov+1))){
    covariate_list[[i]] <- rnorm(sample_size, mean = 0, sd = 1) ## Xs
  }
```

We partition the simulated covariate matrix $X$ into testing, training, validation in propotion 6:2:2 using the `Index_list` created from `creat_testIndex(sample_size)`.
```{r eval=FALSE}
  covariate_training <- list()
  covariate_testing <- list()
  covariate_validation <- list()
  for (i in 1:num_of_cov){
    covariate_training[[i]] <- covariate_list[[i+1]][unlist(Index_list[1:6])]
    covariate_testing[[i]] <- covariate_list[[i+1]][unlist(Index_list[7:8])]
    covariate_validation[[i]] <- covariate_list[[i+1]][unlist(Index_list[9:10])]
  }
```

We simulate a vector of errors from $\mathcal{N}(0,1)$ with the length equal to the sample size.
```{r eval=FALSE}
  e <- rnorm(sample_size, mean = 0, sd = 1)
```

We simulate Y from the the underlying true model $Y = g(X) + \epsilon$ and $g(X) = X\beta$. In this case we only have one covariate $X_1$, so Y comes from $Y = \beta_0 + X_1\beta_1 + \epsilon$.
```{r eval=FALSE}
  y <- Map('*', beta_list, covariate_list) %>% Reduce('+',.) + e
  
  ## if Y has a quatratic term
  #y <- Map('*', beta_list[1:(length(beta_list) - 1)], covariate_list[1:(length(beta_list) - 1)]) %>%
  #     Reduce('+',.) + beta_list[[length(beta_list)]] * covariate_list[[length(beta_list)]]^2 + e
```

We then partition the simulated true Y values into testing, training, validation sets in propotion 6:2:2 using the `Index_list` created from `creat_testIndex(sample_size)`.
```{r eval=FALSE}
  ## partition simulated true Y into testing, training, validation in propotion 6:2:2
  y_training <- y[unlist(Index_list[1:6])]
  y_testing <- y[unlist(Index_list[7:8])]
  y_validation <- y[unlist(Index_list[9:10])]
```

In the training data set, we can fit any model to learn the relationship between the response and covariates. For example, we fit a random forest machine learning model in the training data set, denoted as $f(x)$. 
```{r eval=FALSE}
  ## fit a machine learning model (random forest -- 'method = ranger') using traing data Y ~ X, and get predictions for testing data sets
  training_data <- data.frame(Y = y_training, X = matrix(unlist(covariate_training), nrow = num_of_cov, byrow = TRUE) %>% t())

  model <- train(Y ~., data = training_data,
             method = 'rf',
             importance = TRUE)

```


Then we can obtain the predicted Y value in the testing data set using $f(x)$ by $\hat{Y}_{test} = f(X_{test})$， and the predicted Y value in the validation data set by $\hat{Y}_{val} = f(X_{val})$.
```{r eval=FALSE}
  ## calculate expected y in the testing dataset using model (random forrest)
  testing_covariates <- data.frame(X = matrix(unlist(covariate_testing), nrow = num_of_cov, by = TRUE) %>% t())
  y_hat_testing <- predict(model, testing_covariates)

  validation_covariates <- data.frame(X = matrix(unlist(covariate_validation), nrow = num_of_cov, by = TRUE) %>% t())
  y_hat_validation <- predict(model, validation_covariates)
```

[PLOTS]
```{r eval=FALSE}
## plot y_testing and y_hat_testing in 2D plot (explore joint distribution of y_testing and y_hat_testing)
## joint distribution of y_testing and y_hat_testing is approximately multivariate(bivariate) normal
colors <- densCols(y_testing, y_hat_testing, bandwidth = 2)
data <- data.frame(y_testing, y_hat_testing)
p_yhat_y <- ggplot(data, aes(x = y_testing, y = y_hat_testing)) +
            geom_point(color = colors, size = 2) +
            geom_abline(linetype="dashed",color="red",size=1) +
            labs(y="Predicted Y in testing set (LM)", x = "True Y in testing set")
print(p_yhat_y)
```

The true underlying model $Y = g(X) + \epsilon$ is unknown to us, but we want to fit a linear regression model, $Y = X\beta + \epsilon^*$, on true value $Y$ and covariate matrix $X$. We can then obtain the OLS estimator $\hat{\beta}$ as, 
$$
\hat{\beta} = (X^{T}X)^{-1}X^{T}Y. \tag{1}
$$
However, we are unable to calculate the above unbiased OLS estimator $\hat{\beta}_{val}$ in the validation data set, because the true values $Y_{val}$ in the validation data set are unrevealed to us. Thus, we use the known predicted values $\widehat{Y}_{val}$ instead to write the regression model as $\widehat{Y}_{val} = X_{val}\beta_{est} + \epsilon^{*}$. We can then obtain the OLS estimators $\hat{\beta^{*}}$ as, 
$$
\hat{\beta}_{est} = (X_{val}^{T}X_{val})^{-1}X_{val}^{T}\widehat{Y}_{val}. \tag{2}
$$
```{r eval=FALSE}
  X_val_matrix <- cbind(rep(1,nrow(validation_covariates)),validation_covariates) %>% as.matrix()
  X_test_matrix <- cbind(rep(1,nrow(testing_covariates)),testing_covariates) %>% as.matrix()

  beta_hat_est <- solve(t(X_val_matrix) %*% X_val_matrix) %*% t(X_val_matrix) %*% y_hat_validation
```

We assume that the predicted value $\widehat{Y}$, where $\widehat{Y} = f(X)$, follows a normal distribution centered around the true value $Y$. So, we propose that the mean of $\widehat{Y}$ is a linear function of $Y$. We write out the distribution of $\widehat{Y}$ as, 
$$
\widehat{Y} \sim \mathcal{N}(\gamma_0 + \gamma_1 Y , \sigma_{\widehat{Y}}^2). \tag{3}
$$
Here, we also propose that the mean of true value $Y$ follows a linear regression model $X\beta$, and variance of true value $Y$ is denoted as $\sigma_{Y}^2$. Then we attempt to find the unknown parameter values of $\gamma_0$, $\gamma_1$, and $\sigma_{\widehat{Y}}^2$ that maximize the likelihood function of equation (3). The MLE of parameters can be written as,
$$
\hat{\gamma_0} = \overline{\widehat{Y}} - \hat{\gamma_1} \overline{Y} \tag{4} 
$$
$$
\hat{\gamma_1} = \frac{\sum_{i=1}^{n} (Y_{i} - \overline{Y})(\widehat{Y}_{i} - \overline{\widehat{Y}})} {\sum_{i=1}^{n} (Y_{i} - \overline{Y})^2} \tag{5} 
$$
$$
\hat{\sigma_{\widehat{Y}}^2} = \frac{\sum_{i=1}^{n} (\widehat{Y}_{i} - \hat{\gamma_0} - \hat{\gamma_1} Y_{i})^2} {n} \tag{6}
$$
Since parameters $\gamma_0$, $\gamma_1$, $\beta$ are unknown, we approximate them by using the MLE estimators of $\hat{\gamma_0}_{test}$, $\hat{\gamma_1}_{test}$, and the OLS estimator of $\hat{\beta}_{test}$ calculated in the testing data set,
$$
\hat{\gamma_1}_{test} = \frac{\sum_{i=1}^{n} (Y_{test_i} - \overline{Y}_{test})(\widehat{Y}_{test_i} - \overline{\widehat{Y}}_{test})} {\sum_{i=1}^{n} (Y_{test_i} - \overline{Y}_{test})^2} \tag{7}
$$
```{r eval=FALSE}
  Y_test_mu <- mean(y_testing)
  Y_hat_test_mu <- mean(y_hat_testing)

  gamma1_mle <- sum((y_testing - Y_test_mu) * (y_hat_testing - Y_hat_test_mu)) /
                sum((y_testing - Y_test_mu)^2)
```

$$
\hat{\gamma_0}_{test} = \overline{\widehat{Y}}_{test} - \hat{\gamma_1} \overline{Y}_{test} \tag{8}
$$

```{r eval=FALSE}  
  gamma0_mle <- Y_hat_test_mu - gamma1_mle * Y_test_mu
```

$$
\hat{\beta}_{test} = (X_{test}^{T}X_{test})^{-1}X_{test}^{T} Y_{test} \tag{9}
$$
```{r eval=FALSE} 
  beta_hat_test <- solve(t(X_test_matrix) %*% X_test_matrix) %*% t(X_test_matrix) %*% y_testing
```

[PLOTS]
```{r eval=FALSE}
  ## plot y_hat_validation and est_hat_validation in 2D plot
  est_hat_validation <- gamma0_mle + gamma1_mle * y_validation #X_val_matrix %*% beta_hat_test
  colors <- densCols(y_hat_validation, est_hat_validation, bandwidth = 2)
  data <- data.frame(y_hat_validation, est_hat_validation)
  p_yhat_y <- ggplot(data, aes(x = y_hat_validation, y = est_hat_validation)) +
              geom_point(color = colors, size = 2) +
              geom_abline(linetype="dashed",color="red",size=1) +
              labs(y="estimated Y_hat in validation set (MLE)", x = "Y_hat in validation set")
  print(p_yhat_y)
```

The bias of the OLS estimator $\hat{\beta}$ from equation (1) relative to model assumption $\beta$ is 0,
$$
Bias_{\beta} (\hat{\beta}) = \mathbf{E}_{Y \mid X} (\hat{\beta}) - \beta = 0 \tag{10}
$$
In this case, the bias of $\hat{\beta}_{est}$ from equation (2) relative to $\beta$ must not equal to 0 due to errors introduced by the prediction model $\widehat{Y}_{val} = f(X_{val})$. Therefore, if we estimate such bias using the existing data in the testing data set, and correct this bias in $\hat{\beta}_{est}$ to better estimate the model assumption $\beta$ in the validation data set.
Let us write the bias of $\hat{\beta}_{est}$ relative to $\beta$,
$$
Bias_{\beta} (\hat{\beta}_{est}) = \mathbf{E}_{\widehat{Y} \mid X} (\hat{\beta}_{est}) - \beta \tag{11}
$$
In this case, we further use $\widehat{Y}_{test}$ and X covariate matrix in the testing data set to estimate $\mathbf{E}_{\widehat{Y} \mid X}$, and $Y_{test}$ and X covariate matrix in the testing data set to estimate $\beta$. Thus, we estimate the bias of $\hat{\beta}_{est}$ relative to $\beta$ to be
$$
Bias_{\beta} (\hat{\beta}_{est}) \approx  (X_{test}^{T}X_{test})^{-1}X_{test}^{T} \widehat{Y}_{test}  - \hat{\beta}_{test} \tag{12}
$$
```{r eval=FALSE}
  #Bias1 <- (solve(t(X_val_matrix) %*% X_val_matrix) %*% t(X_val_matrix)) %*%
  #        (gamma0_mle + gamma1_mle * X_val_matrix %*% beta_hat_test) - beta_hat_test
  beta_hat_test_yhat <- solve(t(X_test_matrix) %*% X_test_matrix) %*% t(X_test_matrix) %*% y_hat_testing
  Bias2 <-  beta_hat_test_yhat -  beta_hat_test
```  

Now we propose a new bias-corrected estimator $\hat{\beta}_{BCest}$ to better estimate the true $\beta$ using predicted values $\widehat{Y}_{val}$, where $\hat{\beta}_{BCest}$ is an estimator that corrects the bias of $\hat{\beta}_{est}$ relative to $\beta$ in equation (12).
$$
\hat{\beta}_{BCest} = \hat{\beta}_{est} - Bias_{\beta} (\hat{\beta}_{est}) \tag{13}
$$
```{r eval=FALSE}
  beta_hat_BCest <- beta_hat_est - Bias2
```

Let us calculate the mean of the bias-corrected estimator $\hat{\beta}_{BCest}$ as, 
$$
\begin{split}
& \mathbf{E}_{\widehat{Y} \mid X} (\hat{\beta}_{BCest})\\
& = \mathbf{E}_{\widehat{Y} \mid X} \Big( \hat{\beta}_{est} - Bias_{\beta} (\hat{\beta}_{est}) \Big)\\
& = \mathbf{E}_{\widehat{Y} \mid X} (\hat{\beta}_{est}) - \mathbf{E}_{\widehat{Y} \mid X} (\mathbf{E}_{\widehat{Y} \mid X} (\hat{\beta}_{est} - \beta))\\
& \approx \hat{\beta}_{test}
\end{split}\tag{14}
$$
Now we want to estimate the variance of the bias-corrected estimator $\hat{\beta}_{BCest}$. In order to do so, let us calculate the marginal variance of $\widehat{Y}$ first. Recall from equation (3), we propose that $\widehat{Y} \sim \mathcal{N}(\gamma_0 + \gamma_1 Y , \sigma_{\widehat{Y}}^2)$. Then, we use the MLE estimators from equation (4-6) to calculate the mariginal variance of $\widehat{Y}$,
$$
\begin{split}
\textbf{Var}_{\widehat{Y}} (\widehat{Y}) 
& = \mathbf{E}_{Y} \big(\textbf{Var}_{\widehat{Y} \mid Y} (\widehat{Y} \mid Y) \big) + \textbf{Var}_{Y} \big(\mathbf{E}_{\widehat{Y} \mid Y} (\widehat{Y} \mid Y) \big)\\
& = \sigma_{\widehat{Y}}^2 + \textbf{Var}_{Y} (\gamma_0 + \gamma_1 Y)\\
& = \sigma_{\widehat{Y}}^2 + \gamma_1^2 \textbf{Var}_{Y} (Y)\\
& = \sigma_{\widehat{Y}}^2 + \gamma_1^2 \sigma_{Y}^2
\end{split} \tag{15}
$$
Then we approximate parameters $\gamma_1$ using the MLE estimators of $\hat{\gamma_1}_{test}$ from equation (7).

Let us calculate the mean of the bias-corrected estimator $\hat{\beta}_{BCest}$ as,
$$
\begin{split}
\textbf{Var}_{\widehat{Y} \mid X} (\hat{\beta}_{BCest})
& = \textbf{Var}_{\widehat{Y} \mid X} \Big( \hat{\beta}_{est} - Bias_{\beta} (\hat{\beta}_{est})  \Big)\\
& = \textbf{Var}_{\widehat{Y} \mid X} (\hat{\beta}_{est})\\
& = \textbf{Var}_{\widehat{Y} \mid X} \Big( (X_{val}^{T}X_{val})^{-1}X_{val}^{T}\widehat{Y}_{val} \Big)\\
& = (X_{val}^{T}X_{val})^{-1}X_{val}^{T} \cdot \textbf{Var}_{\widehat{Y} \mid X} (\widehat{Y}_{val}) \cdot X_{val}(X_{val}^{T}X_{val})^{-1}\\
& = (X_{val}^{T}X_{val})^{-1} X_{val}^{T} \cdot (\sigma^2_{\widehat{Y}} + \gamma_1^2 \sigma_{Y}^2) \cdot X_{val}(X_{val}^{T}X_{val})^{-1}\\
& \approx (X_{val}^{T}X_{val})^{-1} (\hat{\sigma^2}_{\widehat{Y}_{test}} + \hat{\gamma_1}_{test}^2 \cdot \sigma_{Y_{test}}^2)
\end{split} \tag{16}
$$
```{r eval=FALSE}
  sigma_Y_hat_test <- (sum((y_hat_testing - gamma0_mle - gamma1_mle * y_testing)^2) / (length(y_testing)-1))
  sigma_Y_test <- sum((y_testing - X_test_matrix %*% beta_hat_test)^2) / (length(y_testing)-1)
  var_BCest <- solve(t(X_val_matrix) %*% X_val_matrix) * (sigma_Y_hat_test + gamma1_mle^2 * sigma_Y_test)
```


We want to calculate t statistics of $\hat{\beta}$ estimators to test null hypothesis $H_0: \beta = \beta_{true}$. First, we need to calcuate $\beta_{true}$ using $g(X)$ in the validation data set. In this simulation process, we set $g(X)$ to be a linear model $g(X) = X\beta$. Recall that $Y = g(X) + \epsilon$. So,
$$
\begin{split}
\beta_{true} 
& = \mathbf{E} \big[ (X_{val}^{T}X_{val})^{-1}X_{val}^{T}Y | g(X_{val}) \big] \\
& = (X_{val}^{T}X_{val})^{-1}X_{val}^{T} \mathbf{E} \big[Y | g(X_{val}) \big] \\
& = (X_{val}^{T}X_{val})^{-1}X_{val}^{T}g(X_{val})
\end{split} \tag{17}
$$
```{r eval=FALSE}
  g_xval <- Map('*', beta_list, c(rep(1,length(covariate_validation[[1]])) %>% list(),covariate_validation)) %>%             Reduce('+',.)
  #g_xval <- beta_list[[1]] + beta_list[[2]] * covariate_validation[[1]]^2
  beta_true <- solve(t(X_val_matrix) %*% X_val_matrix) %*% t(X_val_matrix) %*% g_xval
```

Null Hypothesis: $H_0: \beta = \beta_{true}$
a. calcuate t statistics for $t_{\hat{\beta}_{est}}$ using standard error from only fitting the linear model:
$$
\begin{split}
t_{1 \hat{\beta}_{est}(se_{\hat{\beta}_{est}})} 
& = \frac{\hat{\beta}_{est} - \beta_{true}} {se_{\hat{\beta}_{est}}}\\
& = \frac{\hat{\beta}_{est} - \beta_{true}} {\sqrt{(X_{val}^{T}X_{val})^{-1} (\widehat{Y}_{val} - X_{val}\hat{\beta}_{est})^{T} (\widehat{Y}_{val} - X_{val}\hat{\beta}_{est}) / df}}
\end{split}\tag{18}
$$
```{r eval=FALSE}  
## predicted Y_hat_val with lm standard error
  t_b_est_lmse <- (beta_hat_est - beta_true) / sqrt(diag(solve(t(X_val_matrix) %*% X_val_matrix) * sum((y_hat_validation - X_val_matrix %*% beta_hat_est)^2)/(length(y_validation - 1))))
```

b. calcuate t statistics for $t_{\hat{\beta}_{est}}$ using bias-corrected standard error of $\hat{\beta}_{BCest}$:
$$
\begin{split}
t_{1 \hat{\beta}_{est}(se_{\hat{\beta}_{BCest}})} 
& = \frac{\hat{\beta}_{est} - \beta_{true}} {se_{\hat{\beta}_{BCest}}}\\
& = \frac{\hat{\beta}_{est} - \beta_{true}} {\sqrt{\textbf{Var}_{\widehat{Y} \mid X} (\hat{\beta}_{BCest})}}
\end{split}\tag{19}
$$
```{r eval=FALSE}  
## predicted Y_hat_val with bias corrected standard error
  t_b_est_BCse <- (beta_hat_est - beta_true) / sqrt(diag(var_BCest))
```

c. calcuate t statistics for $t_{\hat{\beta}_{BCest}}$ using standard error of only fitting the linear model:
$$
\begin{split}
t_{1 \hat{\beta}_{BCest}(se_{\hat{\beta}_{est}})} 
& = \frac{\hat{\beta}_{BCest} - \beta_{true}} {se_{\hat{\beta}_{est}}}\\
& = \frac{\hat{\beta}_{BCest} - \beta_{true}} {\sqrt{(X_{val}^{T}X_{val})^{-1} (\widehat{Y}_{val} - X_{val}\hat{\beta}_{est})^{T} (\widehat{Y}_{val} - X_{val}\hat{\beta}_{est}) / df}}
\end{split}\tag{20}
$$
```{r eval=FALSE}  
 ## bias corrected beta with lm standard error
  t_b_BCest_lmse <- (beta_hat_est - Bias2 - beta_true) / sqrt(diag(solve(t(X_val_matrix) %*% X_val_matrix) * sum((y_hat_validation - X_val_matrix %*% beta_hat_est)^2)/(length(y_validation - 1))))
```

d. calcuate t statistics for $t_{\hat{\beta}_{BCest}}$ using bias-corrected standard error of $\hat{\beta}_{BCest}$
$$
\begin{split}
t_{1 \hat{\beta}_{BCest}(se_{\hat{\beta}_{BCest}})} 
& = \frac{\hat{\beta}_{BCest} - \beta_{true}} {se_{\hat{\beta}_{BCest}}}\\
& = \frac{\hat{\beta}_{BCest} - \beta_{true}} {\sqrt{\textbf{Var}_{\widehat{Y} \mid X} (\hat{\beta}_{BCest})}}
\end{split}\tag{21}
$$
```{r eval=FALSE}  
 ## bias corrected beta with bias corrected standard error
  t_b_BCest_BCse <- (beta_hat_est - Bias2 - beta_true) / sqrt(diag(var_BCest))
```
e. calculate t statistics for $t_{\hat{\beta}_{val}}$ using true $Y_{val}$ values in the validation data set. Note that we do not know $Y_{val}$ in real life. We can only calculate this statistics using simulation data with known data generation mechanism.
$$
\begin{split}
t_{1 \hat{\beta}_{val}(se_{\hat{\beta}_{val}})} 
& = \frac{\hat{\beta}_{val} - \beta_{true}} {se_{\hat{\beta}_{val}}}\\
& = \frac{\hat{\beta}_{val} - \beta_{true}} {\sqrt{(X_{val}^{T}X_{val})^{-1} (Y - X_{val}\hat{\beta}_{val})^{T} (Y - X_{val}\hat{\beta}_{val}) / df}}
\end{split}\tag{22}
$$
```{r eval=FALSE}  
  beta_val <- solve(t(X_val_matrix) %*% X_val_matrix) %*% t(X_val_matrix) %*% y_validation
  ## lm using known Y_val
  t_b_val <- (beta_val - beta_true) / sqrt(diag(solve(t(X_val_matrix) %*% X_val_matrix) * sum((y_validation - X_val_matrix %*% beta_val)^2)/(length(y_validation - 1))))
```
 
We also want to calculate t statistics of $\hat{\beta}$ estimators to test null hypothesis $H_0: \beta = 0$.
 
Null Hypothesis: $H_0: \beta = 0$
a. calcuate t statistics for $t_{\hat{\beta}_{est}}$ using standard error from only fitting the linear model:
$$
\begin{split}
t_{2 \hat{\beta}_{est}(se_{\hat{\beta}_{est}})} 
& = \frac{\hat{\beta}_{est} - 0} {se_{\hat{\beta}_{est}}}\\
& = \frac{\hat{\beta}_{est} } {\sqrt{(X_{val}^{T}X_{val})^{-1} (\widehat{Y}_{val} - X_{val}\hat{\beta}_{est})^{T} (\widehat{Y}_{val} - X_{val}\hat{\beta}_{est}) / df}}
\end{split}\tag{23}
$$
```{r eval=FALSE} 
  t_0_est_lmse <- beta_hat_est / sqrt(diag(solve(t(X_val_matrix) %*% X_val_matrix) * sum((y_hat_validation - X_val_matrix %*% beta_hat_est)^2)/(length(y_validation - 1))))
```

b. calcuate t statistics for $t_{\hat{\beta}_{est}}$ using bias-corrected standard error of $\hat{\beta}_{BCest}$:
$$
\begin{split}
t_{2 \hat{\beta}_{est}(se_{\hat{\beta}_{BCest}})} 
& = \frac{\hat{\beta}_{est} -0} {se_{\hat{\beta}_{BCest}}}\\
& = \frac{\hat{\beta}_{est}} {\sqrt{\textbf{Var}_{\widehat{Y} \mid X} (\hat{\beta}_{BCest})}}
\end{split}\tag{24}
$$
```{r eval=FALSE} 
t_0_est_BCse <- beta_hat_est / sqrt(diag(var_BCest))
```

c. calcuate t statistics for $t_{\hat{\beta}_{BCest}}$ using standard error of only fitting the linear model:
$$
\begin{split}
t_{2 \hat{\beta}_{BCest}(se_{\hat{\beta}_{est}})} 
& = \frac{\hat{\beta}_{BCest} -0} {se_{\hat{\beta}_{est}}}\\
& = \frac{\hat{\beta}_{BCest} } {\sqrt{(X_{val}^{T}X_{val})^{-1} (\widehat{Y}_{val} - X_{val}\hat{\beta}_{est})^{T} (\widehat{Y}_{val} - X_{val}\hat{\beta}_{est}) / df}}
\end{split}\tag{25}
$$
```{r eval=FALSE} 
t_0_BCest_lmse <- (beta_hat_est - Bias2) / sqrt(diag(solve(t(X_val_matrix) %*% X_val_matrix) * sum((y_hat_validation - X_val_matrix %*% beta_hat_est)^2)/(length(y_validation - 1))))
```

d. calcuate t statistics for $t_{\hat{\beta}_{BCest}}$ using bias-corrected standard error of $\hat{\beta}_{BCest}$
$$
\begin{split}
t_{2\hat{\beta}_{BCest}(se_{\hat{\beta}_{BCest}})} 
& = \frac{\hat{\beta}_{BCest} - 0} {se_{\hat{\beta}_{BCest}}}\\
& = \frac{\hat{\beta}_{BCest}} {\sqrt{\textbf{Var}_{\widehat{Y} \mid X} (\hat{\beta}_{BCest})}}
\end{split}\tag{26}
$$
```{r eval=FALSE} 
t_0_BCest_BCse <- (beta_hat_est - Bias2) / sqrt(diag(var_BCest))
```

e. calculate t statistics for $t_{\hat{\beta}_{val}}$ using true $Y_{val}$ values in the validation data set. Note that we do not know $Y_{val}$ in real life. We can only calculate this statistics using simulation data with known data generation mechanism.
$$
\begin{split}
t_{2 \hat{\beta}_{val}(se_{\hat{\beta}_{val}})} 
& = \frac{\hat{\beta}_{val} -0} {se_{\hat{\beta}_{val}}}\\
& = \frac{\hat{\beta}_{val} } {\sqrt{(X_{val}^{T}X_{val})^{-1} (Y - X_{val}\hat{\beta}_{val})^{T} (Y - X_{val}\hat{\beta}_{val}) / df}}
\end{split} \tag{27}
$$
```{r eval=FALSE} 
  #### t statistics for testing beta = 0 (in the summary table when fitting a linear model)
  t_0_val <- beta_val / sqrt(diag(solve(t(X_val_matrix) %*% X_val_matrix) * sum((y_validation - X_val_matrix %*% beta_val)^2)/(length(y_validation - 1)))) 
```

Let us also compare the root-mean-square error(RMSE) of the bias-corrected improved estimators $\hat{\beta}_{BCest}$ to the RMSE of the estimators $\hat{\beta}_{est}$, 
$$
RMSE(\hat{\beta}_{BCest}) = \sqrt{\mathbf{E} \big((\hat{\beta}_{BCest} - \beta_{true})^2 \big)} \tag{28}
$$

```{r eval=FALSE} 
  rmse_beta_hat_est <- sqrt(mean((beta_hat_est[-1] - beta_true[-1])^2))
  rmse_beta_hat_BCest <- sqrt(mean((beta_hat_BCest[-1] - beta_true[-1])^2))
```

```{r eval=FALSE}
 # function returns a list of estimators and t statistics
  result <- list(Bias2, beta_hat_est, beta_hat_test, beta_val, beta_hat_test_yhat, beta_true,
                rmse_beta_hat_est, rmse_beta_hat_BCest,
                t_b_val, t_b_est_lmse, t_b_BCest_lmse, t_b_est_BCse, t_b_BCest_BCse,
                t_0_val, t_0_est_lmse, t_0_BCest_lmse, t_0_est_BCse, t_0_BCest_BCse)
  return(result)
}
```

```{r eval=FALSE} 
# Run the simulation
num_of_cov = 1
sample_size = 2000

for (beta2 in seq(1,40,1)){
  beta_list <- list(1,beta2)

  set.seed(2018)
  Index_list <- create_testIndex(sample_size)

  result <- mclapply(c(1:100), simulation, sample_size = sample_size, num_of_cov = num_of_cov, beta_list = beta_list, Index_list = Index_list,mc.cores = 20)
  save(result, file = paste0("BC_app2_nonlinear_rf_beta",beta2,".rda"))

}

q(save = "no")
```
